---
title: "Make target dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Make target dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Purpose

This vignette documents the workflow to create the target dataset_
The [training dataset](make_training_dataset.html) is extracted from the target data set at selected labelled locations.


# Libraries

```{r setup}
library(RiverML)
library(raster)
```

# Loading data

The dataset of target streamlines is loaded with `get_target_streamlines()`.
The target streamlines for nine regions are included in the package and accessed with `target_streamlines_SFE`, for `SFE` region corresponding to the catchment of the South Fork Eel River, California, USA.
The nine regions are: 

| Region name           | Region code |
|:----------------------|:-----------:|
| South Fork Eel        |     SFE     |
| Klamath               |      K      |
| North Coast           |     NC      |
| North Central Coast   |     NCC     |
| South Central Coast   |     SCC     |
| South Coast           |     SC      |
| Sacramento            |     SAC     |
| South East California |    SECA     |
| San Joaquin Tulare    |     SJT     |

## Loading `target_streamlines`

We first define a region of study with the variable `region` and check that `get_target_streamlines()` and the lazy loaded dataset `target_streamlines_SFE` are identical.

```{r get_target_streamlines}
region <- "SFE"
target_streamlines <- get_target_streamlines(region)
data(target_streamlines_SFE)
identical(target_streamlines, target_streamlines_SFE)
target_streamlines
```

## Getting `target_points`

From the `target_streamlines`, we now extract the `target_points` using `get_target_points_df()` which returns a `data.frame` containing the location of the midpoints of all the 200-m stream intervals.

```{r get_target_points}
target_points <- get_target_points_df(target_streamlines)
head(target_points)
```

You will want to save this `target_points` `data.frame` as a separate `.csv` with format `region_all_input.csv`. 
Such a file is required for the next step, high performance computing.

```
fname <- paste0(region,"_all_input.csv")
write.csv(target_points, file.path("path/to/output/", fname))
```

# High Performance Computing

At this stage, the goal is to use High Performance Computing (HPC) to highly parallelize the extraction of the data at the location defined by `target_points`.

## HPC Optimization

While it is possible to request one HPC node per location in `target_points`, this is highly unefficient.
This is because, as one HPC admin told me once, if you do so, most of your time might be spent in reading required data, rather than in actual compute.
To estimate the optimum ratio of nodes to compute iterations, we use `HPC_optim()` which requires that you profile one iteration of the code using for example `profvis` package.

```{r, fig.width = 6}
optim_df <- HPC_optim(
	nrow(target_points), 
	hpc_lim = 1e4, 
	io_time = 20e3, 
	computing_time = 1830, 
	AssocGrpCpuLimit = 258, 
	core_per_task = 1
)
optim_df
```

This indicates that we should submit `r optim_df$n_array` HPC jobs each computing `r optim_df$task_per_array` iterations with an estimated time of `r signif(optim_df$exec_time_min, 3)` minutes per array.
Take this estimated execution time with a grain of salt as the `io_time` and `computing_time` values are usually derived using a local computer with likely more RAM than one HPC node, and with faster I/O speed than the I/O speed between the HPC nodes and wherever is your input data stored.

## Getting the terrain analysis data

Armed with the optimized number of arrays and tasks per arrays, the HPC job can be submitted.
This job will compute terrain analysis metrics (TAM, e.g. slope, curvature), and more specifically metrics of such TAM (e.g., mean, max).
To derive these, a digital elevation model (DEM) is tiled around the locations specified by `target_points`.
In our case, the input DEM corresponds to a 10-m DEM of California and its size prevents use to make the rest of the vignette executable.

In the following, the HPC code is detailled and commented.

### Submitting

The TAM-DM HPC job is submitted with the following chunk which you will need to tailor to your own needs.
The arguments `$1` and `$2` specify the output directory and the type of run.
One example for `$1` is `/home/hguillon/out/run124`.
One example for `$2` is `SFE_all`.

```
#!/bin/bash -l

#SBATCH --partition=med
#SBATCH --time=00-03:00:00
#SBATCH --mem-per-cpu=5200
#SBATCH --mail-type=END
#SBATCH --mail-user=hguillon@ucdavis.edu
#SBATCH --job-name=main_process_tam-dm
#SBATCH --output=/home/hguillon/out/main_process_tam-dm_%A_%a.out
#SBATCH --error=/home/hguillon/out/main_process_tam-dm_%A_%a_err.out

srun Rscript ~/farm_scripts/main_process_tam-dm.R $1 $2
```

### Initialization

We first check for arguments from the submitting `slurm` commands and retrieve the `region`.

```
# checking for args
args = commandArgs(trailingOnly=TRUE)

# test if there is at least one argument: if not, return an error
if (length(args)!=2) {
  stop("Output dir or input file missing.\n", call.=FALSE)
}
outdir <- args[1]
run_type <- args[2]
region <- unlist(strsplit(run_type, "_"))[1]
```


## Getting the StreamCat data